---
title: "Neighborhood-Level Predictors of Late HIV Diagnosis"
---

## Overview

This project examines neighborhood-level socioeconomic predictors of late HIV diagnosis. I served as the lead biostatistician, responsible for analytic strategy, modeling decisions, and interpretation.

> All data shown are simulated or de-identified.

## Data

-   Outcome: Late diagnosis (binary)
-   Covariates: age at diagnosis, race, neighborhood poverty
-   Clustering unit: health district

## Statistical Methods

A multilevel logistic regression model was used to account for clustering of individuals within health districts.

Modeling choices were guided by: - Public health interpretability - Stability under sparse geographic data - Alignment with study objectives


```{r, echo=FALSE}

set.seed(2024)
n <- 1000
dat <- data.frame(
late_dx = rbinom(n, 1, 0.30),
age_dx = rnorm(n, 40, 9),
race = factor(sample(c("White", "Black", "Hispanic", "Other"),
n, replace = TRUE)),
poverty = runif(n, 0.05, 0.45),
district = factor(sample(paste0("D", 1:12), n, replace = TRUE))
)
```

```{r}
library(lme4)

fit <- glmer(
  late_dx ~ age_dx + race + poverty + (1 | district),
  data = dat,
  family = binomial
)
```

## Interpretation

Higher neighborhood poverty was associated with increased odds of late HIV
diagnosis after adjusting for individual-level factors.

Racial disparities persisted, suggesting that neighborhood context alone does
not fully explain observed differences.

## Analytical Decisions & Challenges

- **Geographic unit**: Zip-code–level models were unstable due to sparse counts;
  district-level aggregation improved model convergence.
- **Missing data**: Multiple imputation was used instead of complete-case
  analysis to reduce bias.
- **Model complexity**: Penalized models were explored but not selected due to
  interpretability concerns for stakeholders.

## Impact

- Findings informed geographically targeted HIV testing strategies.
- Results were presented to public health leadership.
- Supported a peer-reviewed manuscript submission.


## Reproducibility

Analyses were conducted in R using Quarto for reproducible reporting and version
control via Git.

