---
title: "R01 Data Visualization"
---

## Overview

R01 visualization will provide the ZIP Code-level HIV/AIDS-related statistics among people with HIV (PWH) who have a current Miami-Dade residential ZIP Code and were diagnosed from 2019 to 2021. 

The main focused outcomes are the proportion of viral non-suppression among PWH who are or are not receiving medical care for HIV. We will additionally infer a genetic network using a genetic distance of 0.5% (a proxy for recent linkage) using HIV all sequences from persons newly diagnosed from 2019 to 2021. Our primary goals are to identify clinical variables (care status, diagnosis stage, and proportion with viral suppression) that is associated with prospective cluster growth. Outcomes will be compared to PWH who do not cluster. The plots will have filters to adjust the genetic distance (0.5%, 1%, and 1.5%), time frame of sampling (the calendar year of 2021 and accumulated years across 2019 – 2021), race, and transmission risk. A comprehensive demographics breakdown will be provided for each group of conditions. 
> All data shown are simulated or de-identified.

## Data

-   Outcome: Proportion of viral non-suppression among PWH who are or are not recieving medical care for HIV in clusters. proportion of not in-care among PLWH in clusters, proportion of Late diagnosis among PWH in clusters. (binary)
-   Subgroup: age at diagnosis, race, Exposure Category, Sequence & cluster vs.Sequence & Uncluster vs. No Sequence 
-   Clustering unit: Zip-code

## Statistical Methods
Analyses are descriptive and visualization-based. ZIP Code–level counts and proportions are summarized and displayed using interactive maps to highlight geographic patterns in HIV outcomes and clustering. Interactive line graphs are used to visualize temporal trends in outcomes stratified by HIV sequence and clustering status. All analyses are conducted in R and RShiny using reproducible workflows.

```{r, echo=FALSE}

set.seed(2024)
n <- 1000
dat <- data.frame(
late_dx = rbinom(n, 1, 0.30),
age_dx = rnorm(n, 40, 9),
race = factor(sample(c("White", "Black", "Hispanic", "Other"),
n, replace = TRUE)),
poverty = runif(n, 0.05, 0.45),
district = factor(sample(paste0("D", 1:12), n, replace = TRUE))
)
```

```{r}
    FUN = function(zip, supp, incare, latediag){
      perc_by_zip <- perc_by_zip %>%
        mutate(clustered_group = ifelse(clustered , "clustered", "unclustered")) %>%
        mutate(clustered_group_factor = factor(clustered_group, levels = c("clustered", "unclustered"), ordered = TRUE))
      
      perc_by_zip <- perc_by_zip %>%
        mutate(zip_cd_3years = ifelse(zip_cd_2021 %in% miami_shp_zip$ZIP, zip_cd_2021,
                                      ifelse(zip_cd_2020 %in% miami_shp_zip$ZIP, zip_cd_2020,
                                             ifelse(zip_cd_2019 %in% miami_shp_zip$ZIP, zip_cd_2019, NA)
                                      )))%>%
        mutate(zip_cd_2021 = as.numeric(zip_cd_2021), zip_cd_3years = as.numeric(zip_cd_3years))
      
      perc_by_zip <- perc_by_zip %>%
        filter( (`Vital Status` %in% c("Alive", NA) | death_year >= 2021)) %>%
        mutate(zip = ifelse(zip %in% miami_shp_zip$ZIP, zip, NA)) 
      
      perc_by_zip_denom <- perc_by_zip %>%
        group_by(!!sym(zip)) %>%
        summarize(DenomPLWH = n())
      
      perc_by_zip <- perc_by_zip %>%
        group_by(!!sym(zip), clustered_group) %>%
       filter(clustered_group == "clustered") %>%
        mutate(supp.incare = ifelse( !!sym(supp) %in% 1 & !!sym(incare) %in% 1, 1, 0))%>%
        summarize(
          # Plots A, B, C, D
          PLWH = n(),
          # Plot B (Viral supression)
          num.supp = sum(!!sym(supp), na.rm = TRUE),
          num.unsupp = PLWH - num.supp,
          pop.supp = sum(!!sym(supp), na.rm = TRUE) / PLWH,
          pop.unsupp = 1-pop.supp,
          # Plot C (in-care)
          num.in_care = sum(!!sym(incare), na.rm = TRUE),
          num.not.in_care = PLWH -num.in_care,
          pop.in_care = sum(!!sym(incare), na.rm = TRUE) / PLWH,
          pop.not.in_care = 1-pop.in_care,
          #Plot B2 - Viral suppression among ppl who received medical care
          num.supp.incare = sum(supp.incare, na.rm = TRUE),
          num.unsupp.incare = PLWH - num.supp.incare,
          pop.supp.incare = sum(supp.incare, na.rm = TRUE) / num.in_care,
          pop.unsupp.incare = 1-pop.supp.incare,
          #Plot D (late diagnosis)
          num.late_diagnosis = sum(!!sym(latediag), na.rm = TRUE),
          pop.late_diagnosis = sum(!!sym(latediag), na.rm = TRUE) / PLWH
        )  %>%
        left_join(perc_by_zip_denom, by = c(zip)) %>%
        mutate(pop.plwh = PLWH/DenomPLWH) %>%
        mutate(across(starts_with("pop"), \(x) round(x, 2)))
      
      md_perc_zips <- miami_shp_zip %>%
        left_join(perc_by_zip, by = c("ZIP" = zip))
    }
```

## Interpretation

The proportions of late HIV diagnosis, not being in care, and viral non-suppression were lower among individuals in the sequenced and clustered group compared with other groups.

## Analytical Decisions & Challenges

- **Geographic unit**: Zip-code–level models were unstable due to sparse counts;
  district-level aggregation improved model convergence.

## Impact

- These findings suggest that individuals in sequenced and clustered groups are more likely to be engaged in HIV care and achieve viral suppression. The results also help identify geographic characteristics and communities where HIV treatment and prevention efforts may be more effective or require additional focus.
